---
title: "vignette-geospatial"
author: "Liuqian Bao, Xiaofeng Cai, Jiajia Feng, Shi Xue, Jiahui He"
date: today 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Spatial Analysis of Travel Distances Across Different Modes Using Geographically Weighted Regression (GWR)

## Goal 
The goal is to understand how travel behavior (e.g., distances traveled for different modes of transportation) varies across different counties through a spatial regression analysis. 

## Data Sources
This vignette uses a database of 26,095 sample households residing in California, containing detailed information on their travel behavior on one assigned day for each household, from April 19, 2016 through April 25, 2017, provided by the Transportation Secure Data Center (TSDC). [1]
Specifically, we obtained travel mode data from `PersonData.Rds`, `HHData.Rds`and geographical data, i.e. the physical location and shape of each county, from`counties.shp`, files given in the dataset.

[1]"Transportation Secure Data Center." (2019). National Renewable Energy Laboratory. Accessed Jan. 15, 2019: www.nrel.gov/tsdc.

## Methodology
### Geographically Weighted Regression (GWR): 

Geographically Weighted Regression (GWR) is a spatial analysis technique that extends traditional regression by allowing the relationships between dependent and independent variables to vary spatially. Unlike ordinary least squares (OLS) regression, which assumes global stationarity of the coefficients, GWR incorporates geographic context into the model. This approach accounts for spatial heterogeneity, a common characteristic in spatial datasets, where relationships can change over space due to localized factors.

In GWR, the regression is performed repeatedly for each location in the dataset, weighting observations according to their spatial proximity to the focal location. The weighting is determined using a kernel function, which can be fixed or adaptive, depending on the data's spatial distribution.[2] 

[2] Charlton, M., & Fotheringham, A. S. (2009). Geographically weighted regression. [White Paper].

## Data Pre-Processing

```{r}
library(sf) 
library(sp)
library(dplyr)
library(GWmodel)
library(ggplot2)
library(mapview)
library(leafpop)

person_data <- readRDS("./data/PersonData.Rds")
hh_data <- readRDS("./data/HHData.Rds")
bg_density <- readRDS("./data/hh_bgDensity.Rds")
shapefile <- st_read("./data/counties/counties.shp")
boundaries <- st_read("./data/ca_state_boundaries/CA_State.shp")

head(person_data)
head(hh_data)
head(bg_density)
head(shapefile)
```


```{r}
# Merge Data
combine_data <- left_join(person_data, hh_data) %>% left_join(bg_density)
head(combine_data)
```


Group the data by county (CTFIP), and calculates the average total number of miles traveled by a person for each travel mode (Drive Alone, Drive with Others, Passenger, Walk, and Total) on the survey day.
```{r}
summarized_data <- combine_data %>%
  select(hhid, pnum, DriveAlone_Dist, Driveothers_Dist, Passenger_Dist, Walk_Dist, Bike_Dist, CTFIP, Sum_PMT) %>%
  group_by(CTFIP) %>%
  summarise(
    avg_DriveAlone_Dist = mean(DriveAlone_Dist, na.rm = T),
    avg_Driveothers_Dist = mean(Driveothers_Dist, na.rm = T),
    avg_Passenger_Dist = mean(Passenger_Dist, na.rm = T),
    avg_Walk_Dist = mean(Walk_Dist, na.rm = T),
    avg_Bike_Dist = mean(Bike_Dist, na.rm = T),
    avg_Sum_Pmt = mean(Sum_PMT, na.rm = T))
```


Plot
```{r}
shapefile_sum <- shapefile %>%
  left_join(summarized_data, by = "CTFIP")

mapviewOptions(fgb = F)

mapview(shapefile_sum, 
        zcol = "avg_Sum_Pmt", # assigned color based on sum distance
        legend = TRUE,
        label = as.character(shapefile_sum$CTFIP),  
        popup = popupTable(shapefile_sum, 
                           zcol = c("avg_DriveAlone_Dist", 
                                    "avg_Driveothers_Dist", 
                                    "avg_Passenger_Dist", 
                                    "avg_Walk_Dist", 
                                    "avg_Bike_Dist",
                                    "avg_Sum_Pmt")))
```


#### Select GWR Bandwidth
Use `bw.gwr()` to find the optimal bandwidth for GWR analysis.

```{r}
coords <- st_coordinates(st_centroid(st_geometry(shapefile_sum)))

gwr_data <- shapefile_sum %>%
  select(avg_DriveAlone_Dist, avg_Driveothers_Dist, avg_Passenger_Dist, avg_Walk_Dist, avg_Bike_Dist, avg_Sum_Pmt) %>%
  st_drop_geometry()

gwr_data <- cbind(gwr_data, coords)

gwr_formula <- avg_Sum_Pmt ~ avg_DriveAlone_Dist + avg_Driveothers_Dist + avg_Passenger_Dist + avg_Walk_Dist + avg_Bike_Dist

# convert to spatial data
coordinates(gwr_data) <- ~X + Y
proj4string(gwr_data) <- CRS("+proj=longlat +datum=WGS84")

# Perform bandwidth selection for GWR
bw <- bw.gwr(
  formula = gwr_formula,
  data = gwr_data,  
  adaptive = T)
```

#### Perform Geographically Weighted Regression

```{r}
gwr_model <- gwr.basic(
  formula = gwr_formula,
  data = gwr_data,
  bw = bw, 
  adaptive = T)

gwr_result <- gwr_model$SDF
summary(gwr_result)
head(gwr_result)

gwr_sf <- st_as_sf(gwr_result)
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_DriveAlone_Dist)) + 
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_DriveAlone_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Driveothers_Dist)) +  
  scale_color_viridis_c() +  
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Driveothers_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Passenger_Dist)) +  
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  scale_color_viridis_c() + 
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Passenger_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Walk_Dist)) +  
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Walk_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Bike_Dist)) +  
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Bike_Dist Coefficient", 
       color = "Coefficient Value")
```

interpret coefficient xxxx