---
title: "vignette-geospatial"
author: "Liuqian Bao, Xiaofeng Cai, Jiajia Feng, Shi Xue, Jiahui He"
date: today 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Spatial Analysis of Travel Distances Across Different Modes Using Geographically Weighted Regression (GWR)

## Goal

The goal is to understand how travel behavior (e.g., distances traveled for different modes of transportation) varies across different counties through a spatial regression analysis.

## Data Sources

This vignette uses a database of 26,095 sample households residing in California, containing detailed information on their travel behavior on one assigned day for each household, from April 19, 2016 through April 25, 2017, provided by the Transportation Secure Data Center (TSDC). [1] Specifically, we obtained travel mode data from `PersonData.Rds`, `HHData.Rds`and geographical data, i.e. the physical location and shape of each county, from`counties.shp`, files given in the dataset.

[1]"Transportation Secure Data Center." (2019). National Renewable Energy Laboratory. Accessed Jan. 15, 2019: www.nrel.gov/tsdc.

## Methodology

### Geographically Weighted Regression (GWR):

Geographically Weighted Regression (GWR) is a spatial analysis technique that extends traditional regression by allowing the relationships between dependent and independent variables to vary spatially. Unlike ordinary least squares (OLS) regression, which assumes global stationarity of the coefficients, GWR incorporates geographic context into the model. This approach accounts for spatial heterogeneity, a common characteristic in spatial datasets, where relationships can change over space due to localized factors.

In GWR, the regression is performed repeatedly for each location in the dataset, weighting observations according to their spatial proximity to the focal location. The weighting is determined using a kernel function, which can be fixed or adaptive, depending on the data's spatial distribution.[2]

#### Bandwidth Selection:

The selection of an appropriate bandwidth is a crucial step for GWR model. Bandwidth is a parameter that governs the spatial extent, over which neighboring observations influence the estimation of local parameters. The bandwidth serves as a key filter determining the degree of localization in the analysis.

A bandwidth that is too narrow may lead to oversensitivity to local variations, potentially capturing noise in the data. On the other hand, too broad bandwidths can result in over smoothed representations , masking subtle spatial patterns. With a proper bandwidth value, we are able to achieve the balance to ensure the GWR model accurately captures the true spatial heterogeneity without being unduly influenced by distant observations.

Adaptive bandwidths offer an effective solution, as they can vary based on the size of each geographical area and that of its neighbors. Thus, the model can select a narrower bandwidth in dense areas, and a larger one for suburban areas. [3]

In our analysis, we determined the optimal bandwidth using the \`**bw.gwr\`** function, which is 35.

[2] Charlton, M., & Fotheringham, A. S. (2009). Geographically weighted regression. [White Paper].

[3] Kiani et al.(2024, February 29). *Mastering geographically weighted regression: Key considerations for building a robust model: Geospatial Health*. Mastering geographically weighted regression: key considerations for building a robust model \| Geospatial Health. <https://www.geospatialhealth.net/gh/article/view/1271/1365>

## Data Pre-Processing

### Merging data

```{r warning=FALSE, message = FALSE}
library(sf) 
library(sp)
library(dplyr)
library(GWmodel)
library(ggplot2)
library(mapview)
library(leafpop)

person_data <- readRDS("./data/PersonData.Rds")
hh_data <- readRDS("./data/HHData.Rds")
bg_density <- readRDS("./data/hh_bgDensity.Rds")
shapefile <- st_read("./data/counties/counties.shp")
boundaries <- st_read("./data/ca_state_boundaries/CA_State.shp")

head(person_data)
head(hh_data)
head(bg_density)
head(shapefile)
```

```{r warning=FALSE, message = FALSE}
# Merge Data
combine_data <- left_join(person_data, hh_data) %>% left_join(bg_density)
head(combine_data)
```

Group the data by county (CTFIP), and calculates the average total number of miles traveled by a person for each travel mode (Drive Alone, Drive with Others, Passenger, Walk, and Total) on the survey day.

```{r warning=FALSE}
summarized_data <- combine_data %>%
  select(hhid, pnum, DriveAlone_Dist, Driveothers_Dist, Passenger_Dist, Walk_Dist, Bike_Dist, CTFIP, Sum_PMT) %>%
  group_by(CTFIP) %>%
  summarise(
    avg_DriveAlone_Dist = mean(DriveAlone_Dist, na.rm = T),
    avg_Driveothers_Dist = mean(Driveothers_Dist, na.rm = T),
    avg_Passenger_Dist = mean(Passenger_Dist, na.rm = T),
    avg_Walk_Dist = mean(Walk_Dist, na.rm = T),
    avg_Bike_Dist = mean(Bike_Dist, na.rm = T),
    avg_Sum_Pmt = mean(Sum_PMT, na.rm = T))
```

### Visualizing County Data on a Map

```{r}
shapefile_sum <- shapefile %>%
  left_join(summarized_data, by = "CTFIP")

mapviewOptions(fgb = F)

mapview(shapefile_sum, 
        zcol = "avg_Sum_Pmt", # assigned color based on sum distance
        legend = TRUE,
        label = as.character(shapefile_sum$CTFIP),  
        popup = popupTable(shapefile_sum, 
                           zcol = c("avg_DriveAlone_Dist", 
                                    "avg_Driveothers_Dist", 
                                    "avg_Passenger_Dist", 
                                    "avg_Walk_Dist", 
                                    "avg_Bike_Dist",
                                    "avg_Sum_Pmt")))
```


## Perform Geographically Weighted Regression

### Convert shape file into spatial data 

Convert the `county.shp` file into spatial data 

```{r}
coords <- st_coordinates(st_centroid(st_geometry(shapefile_sum)))

gwr_data <- shapefile_sum %>%
  select(avg_DriveAlone_Dist, avg_Driveothers_Dist, avg_Passenger_Dist, avg_Walk_Dist, avg_Bike_Dist, avg_Sum_Pmt) %>%
  st_drop_geometry()

gwr_data <- cbind(gwr_data, coords)

gwr_formula <- avg_Sum_Pmt ~ avg_DriveAlone_Dist + avg_Driveothers_Dist + avg_Passenger_Dist + avg_Walk_Dist + avg_Bike_Dist

# convert to spatial data
coordinates(gwr_data) <- ~X + Y
proj4string(gwr_data) <- CRS("+proj=longlat +datum=WGS84")
```

### Select GWR Bandwidth

Use `bw.gwr()` to find the optimal bandwidth for GWR analysis.

```{r message = FALSE, results = 'hide'}
# Perform bandwidth selection for GWR
bw <- bw.gwr(
  formula = gwr_formula,
  data = gwr_data,  
  adaptive = T)
```

## Results 

Here is a summary of the geographically weighted regression model fit:

```{r}
gwr_model <- gwr.basic(
  formula = gwr_formula,
  data = gwr_data,
  bw = bw, 
  adaptive = T)

gwr_result <- gwr_model$SDF
summary(gwr_result)
head(gwr_result)
gwr_sf <- st_as_sf(gwr_result)
```

### Plot the coefficients

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_DriveAlone_Dist)) + 
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_DriveAlone_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Driveothers_Dist)) +  
  scale_color_viridis_c() +  
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Driveothers_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Passenger_Dist)) +  
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  scale_color_viridis_c() + 
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Passenger_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Walk_Dist)) +  
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Walk_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Bike_Dist)) +  
  geom_sf(data = boundaries, fill = NA, color = "black", linewidth = 0.2) +
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Bike_Dist Coefficient", 
       color = "Coefficient Value")
```

### Interprotation of the Coefficients

#### Walking

Coastal areas (especially below 38°N) have significantly higher coefficients (yellow to green), suggesting that walking distance contributes more significantly to changes in total traveling distance.

#### Driving Alone

Around 38°N, 123°W which is the Bay Area, the higher coefficients indicate that the average drive-alone distance contributes more significantly to changes in total traveling distance.

#### Driving with Others

Coefficients are higher in southern and northern regions (green regions), while some coastal areas near 38°N have lower coefficients (dark blue), which suggest a converse relationship to the drive-alone distance.

#### Riding as Passanger

The coefficients are generally low (mostly dark blue) across all regions, particularly in southern areas, suggesting that passenger distance is less influential.

#### Biking

Coefficients are lower in southern regions (dark purple regions), suggesting that bike distance is less influential.

#### Summary

**Driving Alone**: Higher influence in the northern regions.

**Driving with Others**: Higher influence in central and southern regions.

**Walking**: Stronger impacts in coastal regions.

**Cycling**: Stronger impacts in northern and central regions.

**Passenger**: Riding as a passenger is relatively consistent across regions.
