<<<<<<< Updated upstream
# Spatial Analysis of Travel Distances Across Different Modes Using Geographically Weighted Regression (GWR)

## Goal 
The goal is to understand how travel behavior (e.g., distances traveled for different modes of transportation) varies across different counties through a spatial regression analysis 

## Data Sources
xxx

## Methodology
Spatial Regression Analysis:
Use Geographically Weighted Regression (GWR) to explore the spatial variation 
Determine how the impact

### Data Pre-Processing
=======
---
editor_options: 
  markdown: 
    wrap: 72
---

# Capturing Spatial Variation of Population Density and Travel Modes Using GWR

## Goal

Analyze how population density influences different travel modes (e.g.,
walking, biking, driving) across regions using Geographically Weighted
Regression (GWR).

Data Sources: Population Density Data: From hh_bgDensity.csv, includes
population density by region. Travel Mode Data: From the main dataset,
including variables like Walk_trips, Bike_trips, and DriveAlone_trips.

Overview: Population Density's Impact on Travel Modes Goal: To analyze
how population density (from hh_bgDensity.csv) influences travel mode
choices, such as walking, biking, and driving alone.

Data Sources: Population Density Data: From hh_bgDensity.csv, includes
population density by region. Travel Mode Data: From the main dataset,
including variables like Walk_trips, Bike_trips, and DriveAlone_trips.

Methodology: Spatial Regression Analysis: Use Geographically Weighted
Regression (GWR) to explore the spatial variation in the relationship
between population density and travel mode choices. Determine how the
impact of population density on travel modes varies across regions.
Group Comparison: Divide regions into high-density and low-density
groups based on population density. Compare travel mode preferences
(e.g., average walking or driving trips) between these groups.

#### load data

>>>>>>> Stashed changes
```{r}
library(sf) 
library(sp)
library(dplyr)
library(GWmodel)
library(ggplot2)
library(mapview)
library(leafpop)

person_data <- readRDS("../Data/PersonData.Rds")
hh_data <- readRDS("../Data/HHData.Rds")
bg_density <- readRDS("../Data/hh_bgDensity.Rds")
shapefile <- st_read("../Data/counties/counties.shp")

head(person_data)
head(hh_data)
head(bg_density)
head(shapefile)
```


```{r}
# Merge Data
combine_data <- left_join(person_data, hh_data) %>% left_join(bg_density)
head(combine_data)
```

9.  A table showing the average distance traveled by each of the
    following modes: Drive Alone, Passenger, and Walking on each day of
    the week for people with disability. (5 points)

<<<<<<< Updated upstream
Group the data by county (CTFIP), and calculates the average total number of miles traveled by a person for each travel mode (Drive Alone, Drive with Others, Passenger, Walk, and Total) on the survey day.
=======
>>>>>>> Stashed changes
```{r}
summarized_data <- combine_data %>%
  select(hhid, pnum, DriveAlone_Dist, Driveothers_Dist, Passenger_Dist, Walk_Dist, Bike_Dist, CTFIP, Sum_PMT) %>%
  group_by(CTFIP) %>%
  summarise(
    avg_DriveAlone_Dist = mean(DriveAlone_Dist, na.rm = T),
    avg_Driveothers_Dist = mean(Driveothers_Dist, na.rm = T),
    avg_Passenger_Dist = mean(Passenger_Dist, na.rm = T),
    avg_Walk_Dist = mean(Walk_Dist, na.rm = T),
    avg_Bike_Dist = mean(Bike_Dist, na.rm = T),
    avg_Sum_Pmt = mean(Sum_PMT, na.rm = T))
```

Plot

```{r}
shapefile_sum <- shapefile %>%
  left_join(summarized_data, by = "CTFIP")

mapviewOptions(fgb = F)

mapview(shapefile_sum, 
        zcol = "avg_Sum_Pmt", # assigned color based on sum distance
        legend = TRUE,
<<<<<<< Updated upstream
        label = as.character(shapefile_sum$CTFIP),  
        popup = popupTable(shapefile_sum, 
                           zcol = c("avg_DriveAlone_Dist", 
                                    "avg_Driveothers_Dist", 
                                    "avg_Passenger_Dist", 
                                    "avg_Walk_Dist", 
                                    "avg_Bike_Dist",
                                    "avg_Sum_Pmt")))
```


#### Select GWR Bandwidth
=======
        label = as.character(shapefile_with_avg$CTFIP),  
        popup = popupTable(shapefile_with_avg, 
                           zcol = c("avg_DriveAlone_Dist", "avg_Driveothers_Dist", 
                                    "avg_Passenger_Dist", "avg_Walk_Dist")))
```

----- not used yet

#### Ensure

-   `hh_data` contains travel mode data: `Walk_trips`, `Bike_trips`,
    `DriveAlone_trips`.
-   `bg_density` includes `region_id`, `population_density`, and
    geographic coordinates (`longitude`, `latitude`).

#### 2.Aggregate Regional Travel Data

Group data by `region_id` and calculate the mean travel mode statistics.

```{r}
# Summarize travel modes by region
trip_summary <- hh_data %>%
  group_by(region_id) %>%
  summarise(
    mean_walk_trips = mean(Walk_trips, na.rm = TRUE),
    mean_bike_trips = mean(Bike_trips, na.rm = TRUE),
    mean_drive_trips = mean(DriveAlone_trips, na.rm = TRUE)
  )

```

#### 4. Prepare Spatial Data

Convert the combined dataset into a spatial object for GWR analysis
using sf.

```{r}
# Convert to spatial data (sf object)
spatial_data <- st_as_sf(
  density_data,
  coords = c("longitude", "latitude"),  # Replace with actual column names for coordinates
  crs = 4326  # Use WGS 84 coordinate reference system
)

# View spatial data summary
print(spatial_data)


```

#### 5. Select GWR Bandwidth

>>>>>>> Stashed changes
Use `bw.gwr()` to find the optimal bandwidth for GWR analysis.

```{r}
coords <- st_coordinates(st_centroid(st_geometry(shapefile_sum)))

gwr_data <- shapefile_sum %>%
  select(avg_DriveAlone_Dist, avg_Driveothers_Dist, avg_Passenger_Dist, avg_Walk_Dist, avg_Bike_Dist, avg_Sum_Pmt) %>%
  st_drop_geometry()

gwr_data <- cbind(gwr_data, coords)

gwr_formula <- avg_Sum_Pmt ~ avg_DriveAlone_Dist + avg_Driveothers_Dist + avg_Passenger_Dist + avg_Walk_Dist + avg_Bike_Dist

# convert to spatial data
coordinates(gwr_data) <- ~X + Y
proj4string(gwr_data) <- CRS("+proj=longlat +datum=WGS84")

# Perform bandwidth selection for GWR
bw <- bw.gwr(
  formula = gwr_formula,
  data = gwr_data,  
  adaptive = T)
```

<<<<<<< Updated upstream
#### Perform Geographically Weighted Regression
=======
#### 6. Perform Geographically Weighted Regression

Run the GWR model for walking trips as an example. Repeat for other
travel modes if needed.

```{r}
# GWR model for walking trips
gwr_result <- gwr.basic(
  formula = mean_walk_trips ~ population_density,
  data = spatial_data,
  bw = bw_walk,  # Use the selected bandwidth
  adaptive = TRUE  # Enable adaptive bandwidth
)
>>>>>>> Stashed changes

```{r}
gwr_model <- gwr.basic(
  formula = gwr_formula,
  data = gwr_data,
  bw = bw, 
  adaptive = T)

gwr_result <- gwr_model$SDF
summary(gwr_result)
head(gwr_result)

gwr_sf <- st_as_sf(gwr_result)
```

<<<<<<< Updated upstream
=======
#### 7. Extract and Visualize GWR Coefficients

Extract coefficients from the GWR results and add them to the spatial
dataset. Create a heatmap to visualize the influence of population
density on walking trips.

>>>>>>> Stashed changes
```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_DriveAlone_Dist)) + 
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_DriveAlone_Dist Coefficient", 
       color = "Coefficient Value")
```

<<<<<<< Updated upstream
=======
#### 8. Compare High-Density and Low-Density Regions

Divide regions into high-density and low-density groups based on the
median population density, then compare travel modes.

>>>>>>> Stashed changes
```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Driveothers_Dist)) +  
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Driveothers_Dist Coefficient", 
       color = "Coefficient Value")
```
<<<<<<< Updated upstream

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Passenger_Dist)) +  
  scale_color_viridis_c() + 
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Passenger_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Walk_Dist)) +  
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Walk_Dist Coefficient", 
       color = "Coefficient Value")
```
=======

#### 9. Interpret Results

-   **GWR Coefficients:** The heatmap shows how the relationship between
    population density and walking trips varies spatially. High
    coefficients indicate regions where population density has a
    stronger impact on walking.
-   **High-Density vs Low-Density Regions:**
    -   High-density regions are likely to favor walking and biking.
    -   Low-density regions may rely more on driving.
>>>>>>> Stashed changes

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Bike_Dist)) +  
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Bike_Dist Coefficient", 
       color = "Coefficient Value")
```

<<<<<<< Updated upstream
interpret coefficient xxxx
=======
1.  **Additional Predictors:**
    -   Include other urbanization proxies like `housing_density` or
        `median_income`.
2.  **Time-Based Analysis:**
    -   If time data is available, analyze how population density
        influences travel modes during different periods (e.g., weekdays
        vs weekends).
3.  **Other Visualization Options:**
    -   Combine travel modes into a single visualization, using distinct
        colors for walking, biking, and driving preferences.
>>>>>>> Stashed changes
