  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_DriveAlone_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Driveothers_Dist)) +  
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Driveothers_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Passenger_Dist)) +  
  scale_color_viridis_c() + 
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Passenger_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Walk_Dist)) +  
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Walk_Dist Coefficient", 
       color = "Coefficient Value")
```

```{r}
ggplot(data = gwr_sf) +
  geom_sf(aes(color = avg_Bike_Dist)) +  
  scale_color_viridis_c() +  
  theme_minimal() +
  labs(title = "Spatial Variation of avg_Bike_Dist Coefficient", 
       color = "Coefficient Value")
```

interpret coefficient xxxx