# Capturing Spatial Variation of Population Density and Travel Modes Using GWR
## Goal 
Analyze how population density influences different travel modes (e.g., walking, biking, driving) across regions using Geographically Weighted Regression (GWR).


Data Sources:
Population Density Data: From hh_bgDensity.csv, includes population density by region.
Travel Mode Data: From the main dataset, including variables like Walk_trips, Bike_trips, and DriveAlone_trips.

Overview: Population Density's Impact on Travel Modes
Goal:
To analyze how population density (from hh_bgDensity.csv) influences travel mode choices, such as walking, biking, and driving alone.

Data Sources:
Population Density Data: From hh_bgDensity.csv, includes population density by region.
Travel Mode Data: From the main dataset, including variables like Walk_trips, Bike_trips, and DriveAlone_trips.

Methodology:
Spatial Regression Analysis:
Use Geographically Weighted Regression (GWR) to explore the spatial variation in the relationship between population density and travel mode choices.
Determine how the impact of population density on travel modes varies across regions.
Group Comparison:
Divide regions into high-density and low-density groups based on population density.
Compare travel mode preferences (e.g., average walking or driving trips) between these groups.


#### load data
```{r}
library(sf)      
library(dplyr)
library(GWmodel)
library(ggplot2)
library(mapview)
library(leafpop)

person_data <- readRDS("../Data/PersonData.Rds")
hh_data <- readRDS("../Data/HHData.Rds")
bg_density <- readRDS("../Data/hh_bgDensity.Rds")
shapefile <- st_read("../Data/counties/counties.shp")

head(person_data)
head(hh_data)
head(bg_density)
head(shapefile)
```

#### 3. Merge All Data

```{r}
# Merge data
combine_data <- left_join(person_data, hh_data) %>% left_join(bg_density)
head(combine_data)
```


9. A table showing the average distance traveled by each of the following modes: Drive Alone, Passenger, and Walking on each day of the week for people with disability. (5 points)
```{r}
summarized_data <- combine_data %>%
  select(hhid, pnum, DriveAlone_Dist, Driveothers_Dist, Passenger_Dist, Walk_Dist, CTFIP) %>%
  group_by(CTFIP) %>%
  summarise(
    avg_DriveAlone_Dist = mean(DriveAlone_Dist, na.rm = TRUE),
    avg_Driveothers_Dist = mean(Driveothers_Dist, na.rm = TRUE),
    avg_Passenger_Dist = mean(Passenger_Dist, na.rm = TRUE),
    avg_Walk_Dist = mean(Walk_Dist, na.rm = TRUE)
  )
```


Plot
```{r}
shapefile_with_avg <- shapefile %>%
  left_join(summarized_data, by = "CTFIP")

mapviewOptions(fgb = F)

mapview(shapefile_with_avg, 
        zcol = "avg_DriveAlone_Dist", 
        legend = TRUE,
        label = as.character(shapefile_with_avg$CTFIP),  
        popup = popupTable(shapefile_with_avg, 
                           zcol = c("avg_DriveAlone_Dist", "avg_Driveothers_Dist", 
                                    "avg_Passenger_Dist", "avg_Walk_Dist")))
```

----- not used yet

#### Ensure
- `hh_data` contains travel mode data: `Walk_trips`, `Bike_trips`, `DriveAlone_trips`.
- `bg_density` includes `region_id`, `population_density`, and geographic coordinates (`longitude`, `latitude`).



#### 2.Aggregate Regional Travel Data
Group data by `region_id` and calculate the mean travel mode statistics.
```{r}
# Summarize travel modes by region
trip_summary <- hh_data %>%
  group_by(region_id) %>%
  summarise(
    mean_walk_trips = mean(Walk_trips, na.rm = TRUE),
    mean_bike_trips = mean(Bike_trips, na.rm = TRUE),
    mean_drive_trips = mean(DriveAlone_trips, na.rm = TRUE)
  )

```



#### 4. Prepare Spatial Data
Convert the combined dataset into a spatial object for GWR analysis using sf.
```{r}
# Convert to spatial data (sf object)
spatial_data <- st_as_sf(
  density_data,
  coords = c("longitude", "latitude"),  # Replace with actual column names for coordinates
  crs = 4326  # Use WGS 84 coordinate reference system
)

# View spatial data summary
print(spatial_data)


```

#### 5. Select GWR Bandwidth
Use `bw.gwr()` to find the optimal bandwidth for GWR analysis.

```{r}
# Determine optimal bandwidth for walking trips
bw_walk <- bw.gwr(mean_walk_trips ~ population_density, data = spatial_data, adaptive = TRUE)
print(bw_walk)


```

#### 6. Perform Geographically Weighted Regression
Run the GWR model for walking trips as an example. Repeat for other travel modes if needed.
```{r}
# GWR model for walking trips
gwr_result <- gwr.basic(
  formula = mean_walk_trips ~ population_density,
  data = spatial_data,
  bw = bw_walk,  # Use the selected bandwidth
  adaptive = TRUE  # Enable adaptive bandwidth
)

# View GWR summary
summary(gwr_result)


```


#### 7. Extract and Visualize GWR Coefficients
Extract coefficients from the GWR results and add them to the spatial dataset. Create a heatmap to visualize the influence of population density on walking trips.
```{r}
# Add GWR coefficients to spatial data
spatial_data$gwr_coeff_walk <- gwr_result$SDF$population_density

# Plot GWR coefficients
ggplot(spatial_data) +
  geom_sf(aes(fill = gwr_coeff_walk)) +
  scale_fill_viridis_c(option = "plasma") +
  labs(
    title = "GWR Coefficients for Population Density (Walking Trips)",
    fill = "Coefficient"
  ) +
  theme_minimal()

```

#### 8. Compare High-Density and Low-Density Regions
Divide regions into high-density and low-density groups based on the median population density, then compare travel modes.
```{r}
# Categorize regions into high-density and low-density groups
density_data <- density_data %>%
  mutate(density_group = ifelse(population_density > median(population_density), "High Density", "Low Density"))

# Summarize average travel modes by density group
group_summary <- density_data %>%
  group_by(density_group) %>%
  summarise(
    avg_walk_trips = mean(mean_walk_trips, na.rm = TRUE),
    avg_bike_trips = mean(mean_bike_trips, na.rm = TRUE),
    avg_drive_trips = mean(mean_drive_trips, na.rm = TRUE)
  )

print(group_summary)

```
#### 9. Interpret Results

- **GWR Coefficients:** The heatmap shows how the relationship between population density and walking trips varies spatially. High coefficients indicate regions where population density has a stronger impact on walking.
- **High-Density vs Low-Density Regions:**
  - High-density regions are likely to favor walking and biking.
  - Low-density regions may rely more on driving.


### **Extensions**

1. **Additional Predictors:**
   - Include other urbanization proxies like `housing_density` or `median_income`.
2. **Time-Based Analysis:**
   - If time data is available, analyze how population density influences travel modes during different periods (e.g., weekdays vs weekends).
3. **Other Visualization Options:**
   - Combine travel modes into a single visualization, using distinct colors for walking, biking, and driving preferences.
